---
title: "GIS"
author: "Sarah Hyland"
format: html
editor: visual
self-contained: true
---

# load relevant packages

```{r}
library(rinat)
library(sf)
library(ggplot2)
library(rosm)
library(ggspatial)
library(leaflet)
library(htmltools)
```

# call all data from iNaturalist

## call Hippocampus capensis data from iNaturalist (2018)

```{r}
obs2018 <- get_inat_obs(taxon_name = "Hippocampus capensis",
                   bounds = c(-34.3, 22.8,-33.8, 23.3),
                   year = 2018,
                   maxresults = 1000)
```

## call Hippocampus capensis data from iNaturalist (2019)

```{r}
obs2019 <- get_inat_obs(taxon_name = "Hippocampus capensis",
                   bounds = c(-34.3, 22.8,-33.8, 23.3),
                   year = 2019,
                   maxresults = 1000)
```

## call Hippocampus capensis data from iNaturalist (2020)

```{r}
obs2020 <- get_inat_obs(taxon_name = "Hippocampus capensis",
                   bounds = c(-34.3, 22.8,-33.8, 23.3),
                   year = 2020,
                   maxresults = 1000)
```

## call Hippocampus capensis data from iNaturalist (2021)

```{r}
obs2021 <- get_inat_obs(taxon_name = "Hippocampus capensis",
                   bounds = c(-34.3, 22.8,-33.8, 23.3),
                   year = 2021,
                   maxresults = 1000)
```

## call Hippocampus capensis data from iNaturalist (2022)

```{r}
obs2022 <- get_inat_obs(taxon_name = "Hippocampus capensis",
                   bounds = c(-34.3, 22.8,-33.8, 23.3),
                   year = 2022,
                   maxresults = 1000)
```

## call Hippocampus capensis data from iNaturalist (2023)

```{r}
obs2023 <- get_inat_obs(taxon_name = "Hippocampus capensis",
                   bounds = c(-34.3, 22.8,-33.8, 23.3),
                   year = 2023,
                   maxresults = 1000)
```

# Group years to make 2 data frames

## group years 2018 - 2020

```{r}
obs1 <- rbind(
  transform(obs2018, year = 2018),
  transform(obs2019, year = 2019),
  transform(obs2020, year = 2020)
)

```

## group years 2021 - 2023

```{r}
obs2 <- rbind(
  transform(obs2021, year = 2021),
  transform(obs2022, year = 2022),
  transform(obs2023, year = 2023)
)

```

# convert dataframes into spatial objects

## convert a dataframe into a spatial object (obs1)

```{r}
obs1 <- st_as_sf(obs1, coords = c("longitude", "latitude"), crs = 4326)
```

## convert dataframe into a spatial object (obs2)

```{r}
obs2 <- st_as_sf(obs2, coords = c("longitude", "latitude"), crs = 4326)
```

# create interactive map

```{r}
leaflet() %>%
  # Add default OpenStreetMap map tiles
  addTiles(group = "Default") %>%  
  # Add our points obs 1
  addCircleMarkers(data = obs1,
                   group = "Hippocampus capensis",
                   radius = 3, 
                   color = "green") %>%
  
  # add our points obs 2
  addCircleMarkers(data = obs2,
                   group = "Hippocampus capensis",
                   radius = 3, 
                  color = "red") %>%
  # add legend
  addLegend("bottomright",
            colors = c("green", "red"),
            labels = c("Years 2018 - 2020", "Years 2020 - 2023"),
            title = "Observations",
            opacity = 1) 
  
```
